---
title: "bugs_love_summers"
output:
  html_document:
    df_print: paged
---



```{r echo=FALSE}
# data import libraries
library(readr)
library(readxl)

# data wrangling libraries
library(tidyr)
library(dplyr)
library(stringr)

# plotting libraries
library(ggplot2)

# mapping libraries
library(ggmap)
library(maps)
library(mapdata)



```

```{r}
# load the dataset
load(file = "data/data.RData")
```


```{r}
# import data 
mosquito <- read_csv("data/aegypti_albopictus.csv", 
    col_types = cols(VECTOR = col_factor(levels = c("Aedes aegypti", 
        "Aedes albopictus"))))

Lyme <- read_csv("data/Lyme_ld-Case-Counts-by-County-00-16.csv")
FIPS <- read_excel("data/US_FIPS_CountyCodes.xls")


# view the file
# head(mosquito, n=3)

```
## Original column names

* VECTOR: Identifying the species; Ae. aegypti or Ae. albopictus 
* OCCURRENCE_ID: Unique identiﬁer for each occurrence in the database after temporal and spatial standardisation. 
* SOURCE_TYPE: Published literature or unpublished sources with reference ID that corresponds to the full list of references in Data Citation 1. 
* LOCATION_TYPE: Whether the record represents a point or a polygon location. 
* POLYGON_ADMIN: Admin level or polygon size which the record represents when the location type is a polygon. −999 when the location type is a point (5km×5km). 
* X: The longitudinal coordinate of the point or polygon centroid (WGS1984 Datum). 
* Y: The latitudinal coordinate of the point or polygon centroid (WGS1984 Datum). 
* YEAR: The year of the occurrence. 
* COUNTRY: The name of the country within which the occurrence lies. 
* COUNTRY_ID: ISO alpha-3 country codes. 
* GAUL_AD0: The country-level global administrative unit layer (GAUL) code (see http://www.fao. org/geonetwork) which identiﬁes the Admin-0 polygon within which any smaller polygons and points lie. 
* STATUS: Established versus transient populations.

```{r echo=FALSE}
# renaming columns to friendlier names:

names(mosquito)[1:12] <-  c(
"bug",
"occurrence_id", 
"source_type",
"location_type",
"polygon_admin", 
"long",
"lat",
"year",
"country",
"country_id",
"gaul_ad0",
"status")

head(mosquito, n=5)

```

```{r}
heating_cooling_degree_days <- read_csv("data/Heating_and_Cooling_Degree_Days_EPA.csv", 
    col_names = c("year","heating_days","cooling_days"), skip = 7)
```

```{r}
GISS <- read_csv("data/Global-mean monthly_seasonal_and annual means_1880-present_GISS.csv", 
    skip = 2)

names(GISS)[names(GISS) == "J-D"] <- "Jan_Dec"
names(GISS)[names(GISS) == "D-N"] <- "Dec_Nov"
names(GISS)
```

```{r}

ggplot(data = heating_cooling_degree_days,
       mapping = aes(x=year, y=cooling_days)) +
  geom_point(color= "red", shape = 1, alpha = 0.5) +
  geom_smooth(color="red") +
  theme_bw()
  
```

```{r}
ggplot(data = heating_cooling_degree_days,
       mapping = aes(x=year, y=heating_days)) +
  geom_point(color = "blue", shape = 1, alpha = 0.5) +
  geom_smooth(color="blue") +
  theme_bw()
```

```{r}

```

```{r echo=FALSE}
mosquito %>% 
  filter(country_id == "USA") %>% 
  ggplot(aes(x=year, fill=bug)) +
    geom_histogram(stat = "bin",  position = "stack")
```

```{r echo=FALSE}
mosquito %>% 
  filter(country == "India") %>% 
  ggplot(aes(x=year, fill=bug)) +
    geom_histogram(stat = "bin",  position = "stack")
```

```{r}
ggplot(data = GISS, mapping = aes(x=Year, y=Jan_Dec)) +
  geom_point(shape = 1, alpha = 0.5) +
  geom_smooth(color = "black") +
  theme_bw()
```

```{r}
# Change column names:
{
  colnames(FIPS)[which(colnames(FIPS) == 'FIPS Code')] <- 'FIPS'
  colnames(FIPS)[which(colnames(FIPS) == 'County Name')] <- 'County'
  
  colnames(Lyme)[which(colnames(Lyme) == "Ctyname")] <- "County"
  colnames(Lyme)[which(colnames(Lyme) == "Stname")] <- "State"
}

# Remove County from counites in Lyme df
Lyme$County <- str_replace_all(Lyme$County," County", "")

# isolate county code 999 as states
{
  Lyme_state <- Lyme %>% 
    filter(CTYCODE == 999) %>% 
    as.data.frame()
}

# isolate counties from Lyme data:
{
  Lyme_counties <- Lyme %>% 
    filter(CTYCODE != 999) %>% 
    as.data.frame()
}

{
# modify the CTYCODE to have 3 zeroes
Lyme_counties$CTYCODE <- str_pad(Lyme_counties$CTYCODE, 3, pad = "0")

# modify the STCODE to have 2 zeoroes
Lyme_counties$STCODE <- str_pad(Lyme_counties$STCODE, 2, pad = "0")

# merge columns
Lyme_counties <- Lyme_counties %>%
  unite(FIPS, c("STCODE","CTYCODE"), sep = "", remove = FALSE)
}
```

```{r}
# Find data different between two county rows
Lyme_counties$County[!FIPS$County%in%Lyme_counties$County]
```

```{r}
# Isolate counties from map_data
counties <- map_data("county") 

# Isolate Lyme incidence data for 2016
Lyme_counties_2016 <- Lyme_counties %>% 
    select(State, County, FIPS, Cases2016)

# rename to match with county map file
names(Lyme_counties_2016) <- c("region", "subregion", "FIPS", "Cases2016")
  
# change to lower case
Lyme_counties_2016$region <-  tolower(Lyme_counties_2016$region)
Lyme_counties_2016$subregion <-  tolower(Lyme_counties_2016$subregion)
  
# merge datasets
lyme_2016_map_data <- inner_join(counties, Lyme_counties_2016, by = c("region", "subregion"))

# # custom ggplot theme to drop the axes and ticks but leave the guides and legends
#   ditch_the_axes <- theme(
#     axis.text = element_blank(),
#     axis.line = element_blank(),
#     axis.ticks = element_blank(),
#     panel.border = element_blank(),
#     panel.grid = element_blank(),
#     axis.title = element_blank()
#   )
  
# test counties map

base_county_map <- ggplot(data = Lyme_counties_2016) + 
    geom_polygon(aes(x = long, y = lat, group = group), fill = "white",color = "gray") + 
    coord_fixed(1.3) +
    guides(fill=FALSE)

lyme_2016_map <- base_county_map + 
    geom_polygon(data = Lyme_counties_2016, aes(fill = Cases2016)) +
    geom_polygon(fill = NA) +
    theme_bw()

lyme_2016_map

# + ditch_the_axes
  
```

```{r}
require(maps)
require(ggplot2)
library(readr)
library(dplyr)
library(RColorBrewer)

# read the incidence rate file:
Lyme_State <- read_delim("data/Lyme_incidence_rate_State.txt", 
                         "\t", escape_double = FALSE, 
                         col_types = cols(`2011` = col_double(),
                                          `2012` = col_double()), 
                         trim_ws = TRUE)

# delete the last row
Lyme_State <- Lyme_State[-nrow(Lyme_State),] 
# change state to region
names(Lyme_State)[1]<-"region"
# View(Lyme_State)

# extract state map data
us_state_map <- map_data('state')
# head(us_state_map)

# extract state lyme data for 2016
year <- "2016"

Lyme_map_data <-  Lyme_State %>% 
  select(region, cases = year) %>% 
  mutate(region = tolower(region))

# merge datasets 
Lyme_map_data <- 
  merge(us_state_map, Lyme_map_data, by='region', all=T)

# plot
ggplot(Lyme_map_data, aes(x=long, y=lat, group=group, fill=cases)) +
  scale_fill_gradientn("",colours=brewer.pal(n=4, "YlOrRd"),na.value = "transparent")+
  geom_polygon()+coord_map()+
  labs(title=paste(year,"Lyme Incidence"),fill = "incidence")+
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1))
  

```

```{r}
# trying a for loop

for (year in c("2006","2010","2015")){
  Lyme_map_data <-  Lyme_State %>% 
    select(region, cases = year) %>% 
    mutate(region = tolower(region))
  
  Lyme_map_data <- 
    merge(us_state_map, Lyme_map_data, by='region', all=T)
  
  gg <- ggplot(Lyme_map_data, aes(x=long, y=lat, group=group, fill=cases)) +
    scale_fill_gradientn("",colours=brewer.pal(n=4, "YlOrRd"))+
    geom_polygon()+coord_map()+
    labs(title=paste(year,"Lyme Incidence"))+
    theme_bw()
  
}
```

```{r}
library(readr)
library(readxl)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)

file1 <- "data/Lyme_ld-Case-Counts-by-County-00-16.csv"
file2 <- "data/US_FIPS_CountyCodes.xls"
Lyme <- read_csv(file1)
FIPS <- read_excel(file2)

# Change column names:
{
  colnames(FIPS)[which(colnames(FIPS) == 'FIPS Code')] <- 'FIPS'
  colnames(FIPS)[which(colnames(FIPS) == 'County Name')] <- 'County'
  colnames(Lyme)[which(colnames(Lyme) == "Ctyname")] <- "County"
  colnames(Lyme)[which(colnames(Lyme) == "Stname")] <- "State"
}
# Remove County from counites in Lyme df
Lyme$County <- str_replace_all(Lyme$County," County", "")

# isolate counties from Lyme data:
{
  Lyme_counties <- Lyme %>%
    filter(CTYCODE != 999) %>%
    as.data.frame()
}

# pad required zeroes for padding up 
Lyme_counties$CTYCODE <- str_pad(Lyme_counties$CTYCODE, 3, pad = "0")

# merge columns
Lyme_counties <- Lyme_counties %>%
  unite(FIPS, c("STCODE","CTYCODE"), sep = "", remove = FALSE)

Lyme_counties$FIPS <- as.numeric(Lyme_counties$FIPS)

# extract data to plot
Lyme_counties_2016 <- Lyme_counties %>%
  select(State, County, FIPS, Cases2016)

# rename to match with county map file
names(Lyme_counties_2016) <- c("region", "subregion", "FIPS", "Cases2016")

# change to lower case
Lyme_counties_2016$region <-  tolower(Lyme_counties_2016$region)
Lyme_counties_2016$subregion <-  tolower(Lyme_counties_2016$subregion)

# ----
# mapping attempt 2
library(rgdal)
library(data.table)

US.counties <- readOGR(dsn="data/gz_2010_us_050_00_5m",layer="gz_2010_us_050_00_5m")

#leave out AK, HI, and PR (state FIPS: 02, 15, and 72)
US.counties <- US.counties[!(US.counties$STATE %in% c("02","15","72")),]
county.data <- US.counties@data
county.data <- cbind(id=rownames(county.data),county.data)
county.data <- data.table(county.data)
county.data[,FIPS:=paste0(STATE,COUNTY)] # this is the state + county FIPS code
setkey(county.data,FIPS)

Lyme_counties_2016$FIPS <- as.character(Lyme_counties_2016$FIPS)
Lyme.data <- data.table(Lyme_counties_2016)
setkey(Lyme.data,FIPS)

county.data[Lyme.data,cases:=Cases2016]
map.df <- data.table(fortify(US.counties))

setkey(map.df,id)
setkey(county.data,id)

map.df[county.data,cases:=cases]

library(RColorBrewer)
ggplot(map.df, aes(x=long, y=lat, group=group, fill=cases)) +
  scale_fill_gradientn(colours = rev(heat.colors(5, alpha=1)), 
                       na.value = "lightgreen", trans = "log10")+
  geom_polygon()+coord_map()+
  labs(title="2016 Lyme cases",x="",y="")+
  theme_bw()
# -----

```


```

