---
title: "bugs_love_summers"
output:
  html_document:
    df_print: paged
---



```{r echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)

```


```{r}
# import data 
mosquito <- read_csv("data/aegypti_albopictus.csv", 
    col_types = cols(VECTOR = col_factor(levels = c("Aedes aegypti", 
        "Aedes albopictus"))))

# view the file
# head(mosquito, n=3)

```
## Original column names

* VECTOR: Identifying the species; Ae. aegypti or Ae. albopictus 
* OCCURRENCE_ID: Unique identiﬁer for each occurrence in the database after temporal and spatial standardisation. 
* SOURCE_TYPE: Published literature or unpublished sources with reference ID that corresponds to the full list of references in Data Citation 1. 
* LOCATION_TYPE: Whether the record represents a point or a polygon location. 
* POLYGON_ADMIN: Admin level or polygon size which the record represents when the location type is a polygon. −999 when the location type is a point (5km×5km). 
* X: The longitudinal coordinate of the point or polygon centroid (WGS1984 Datum). 
* Y: The latitudinal coordinate of the point or polygon centroid (WGS1984 Datum). 
* YEAR: The year of the occurrence. 
* COUNTRY: The name of the country within which the occurrence lies. 
* COUNTRY_ID: ISO alpha-3 country codes. 
* GAUL_AD0: The country-level global administrative unit layer (GAUL) code (see http://www.fao. org/geonetwork) which identiﬁes the Admin-0 polygon within which any smaller polygons and points lie. 
* STATUS: Established versus transient populations.

```{r echo=FALSE}
# renaming columns to friendlier names:

names(mosquito)[1:12] <-  c(
"bug",
"occurrence_id", 
"source_type",
"location_type",
"polygon_admin", 
"long",
"lat",
"year",
"country",
"country_id",
"gaul_ad0",
"status")

head(mosquito, n=5)

```

```{r echo=FALSE}
mosquito %>% 
  filter(country_id == "USA") %>% 
  ggplot(aes(x=year, fill=bug)) +
    geom_histogram(stat = "bin",  position = "stack")
```

```{r echo=FALSE}
mosquito %>% 
  filter(country == "India") %>% 
  ggplot(aes(x=year, fill=bug)) +
    geom_histogram(stat = "bin",  position = "stack")
```

